---
title: "Unempoyment rate for California between 1976 and 2022"
author: "Xintong Chen"
date: "2022-11-05"
output: html_document
---

### Group Member: Xinotng (Ruby) Chen

```{r}
library(tseries)
library(forecast)
library(ggfortify)
library(ggplot2)
library(lmtest)
library(ggpubr)
library(knitr)
library(stargazer)
library(dplyr)
library(ggpubr)
library(TSstudio)
library(urca)
```

```{r}
set.seed(521)
data <- read.csv('CAUR.csv')
data[,1:2]
```

```{r}
summary(data)
```

```{r}
sd(data$CAUR)
```

```{r}
data.ts <- ts(data$CAUR,start=c(1976, 1), frequency=12)
```

```{r}
start(data.ts)
end(data.ts)
frequency(data.ts)
```

```{r}
plot(data.ts)
```

```{r}
# plot ACF
ggAcf(data.ts)
```

```{r}
# plot PACF
ggPacf(data.ts)
```

```{r}
original_data <- data[1:529,]
original_data
```

```{r}
original_ts <- ts(original_data$CAUR,start=c(1976, 1), frequency=12)
start(original_ts)
end(original_ts)
frequency(original_ts)
```

```{r}
autoplot(original_ts) + ylab("Unemployment Rate") + xlab("Years")
```



```{r,echo=FALSE,fig.width=8,fig.height=6,message=FALSE}
# plot ACF
ggAcf(original_ts)
ggsave(
  filename = "original_ts_acf.png", # 保存的文件名称。通过后缀来决定生成什么格式的图片
  width = 8,             # 宽
  height = 6,            # 高
  units = "in",          # 单位
  dpi = 300              # 分辨率DPI
)
```

```{r,echo=FALSE,fig.width=8,fig.height=6,message=FALSE}
# plot PACF
ggPacf(original_ts)
ggsave(
  filename = "original_ts_pacf.png", # 保存的文件名称。通过后缀来决定生成什么格式的图片
  width = 8,             # 宽
  height = 6,            # 高
  units = "in",          # 单位
  dpi = 300              # 分辨率DPI
)
```

Remove Seasonality

```{r}
adf.test(original_ts,k=8)
```

ACF plot decays to zero gradualy, which means the data is non-stationary.

the data has a trend, seasonal component, apparant shape change(we delete the data -- 2020)

```{r}
firstdiff <- diff(original_data[,"CAUR"], diff = 1)
```

```{r, echo=FALSE,fig.width=8,fig.height=6,message=FALSE}
#ACF
ggAcf(firstdiff)
ggsave(
  filename = "firstdiff_ts_acf.png",
  width = 8,             # 宽
  height = 6,            # 高
  units = "in",          # 单位
  dpi = 300              # 分辨率DPI
)
```

```{r, echo=FALSE,fig.width=8,fig.height=6,message=FALSE}
#ACF
ggPacf(firstdiff)
ggsave(
  filename = "firstdiff_ts_pacf.png",
  width = 8,             # 宽
  height = 6,            # 高
  units = "in",          # 单位
  dpi = 300              # 分辨率DPI
)
```

```{r}
adf.test(firstdiff,k=8)
```

```{r}
seconddiff <- diff(firstdiff, diff = 1)
```

```{r, echo=FALSE,fig.width=8,fig.height=6,message=FALSE}
#ACF
ggAcf(seconddiff)
ggsave(
  filename = "seconddiff_ts_acf.png",
  width = 8,             # 宽
  height = 6,            # 高
  units = "in",          # 单位
  dpi = 300              # 分辨率DPI
)
```

```{r, echo=FALSE,fig.width=8,fig.height=6,message=FALSE}
#ACF
ggPacf(seconddiff)
ggsave(
  filename = "seconddiff_ts_pacf.png",
  width = 8,             # 宽
  height = 6,            # 高
  units = "in",          # 单位
  dpi = 300              # 分辨率DPI
)
```

```{r}
adf.test(seconddiff,k=8)
```

```{r, echo=FALSE,fig.width=8,fig.height=6,message=FALSE}
# use differencing to remove seasonal component
# difference(t) = observation(t) - observation(t-1)
original_ts %>% diff() %>% ggtsdisplay()
```

```{r, echo=FALSE,fig.width=8,fig.height=6,message=FALSE}
# use differencing to remove seasonal component
# difference(t) = observation(t) - observation(t-1)
original_ts %>% diff() %>% diff() %>% ggtsdisplay()
```

According to ACF: non-seasonality q = 2, seasonality q = 1
According to PACF: non-seasonality p = 1, seasonality p = 2
difference: non-seasonality d = 2, seasonality p = 0

the data is stationary after differencing.

Using ARIMA(1,1,2)(1,0,0)[12]

```{r}
original_ts %>%
  Arima(order=c(1,2,2),seasonal=c(1,0,0)) %>%
  residuals() %>%
  ggtsdisplay()
```

```{r}
original_ts %>%
  Arima(order=c(1,1,2),seasonal=c(1,0,0)) %>%
  residuals() %>%
  ggtsdisplay()
```

```{r}
original_ts %>%
  Arima(order=c(2,0,0),seasonal=c(1,1,0)) %>%
  residuals() %>%
  ggtsdisplay()
```

```{r}
fit1 <- Arima(original_ts, order=c(1,2,2),seasonal=c(1,0,0))
summary(fit1)
coeftest(fit1)
confint(fit1)
```

```{r,echo=FALSE}
## Check autocorrelation in residuals
Box.test(fit1$resid, type="Ljung-Box")
```

```{r}
fit2 <- Arima(original_ts, order=c(2,1,0),seasonal=c(1,0,0))
summary(fit2)
coeftest(fit2)
confint(fit2)
```

```{r,echo=FALSE}
## Check autocorrelation in residuals
Box.test(fit2$resid, type="Ljung-Box")
```



```{r}
fit2 %>% forecast(h=24) %>% autoplot()
```


```{r}
model_auto <- auto.arima(original_ts)
summary(model_auto)
coeftest(model_auto)
confint(model_auto)
```

```{r,echo=FALSE}
## Check autocorrelation in residuals
Box.test(model_auto$resid, type="Ljung-Box")
```


```{r}
model_auto %>% forecast(h=24) %>% autoplot
```


```{r}
testing <- data[530:561,]
test_ts <- ts(testing$CAUR,start=c(2020, 2), frequency=12)
start(test_ts)
end(test_ts)
frequency(test_ts)
```

```{r}
#In-sample Forecast
AutoArimaForecast <- forecast(model_auto,level=c(95),h=24)
test_forecast(actual=original,forecast.obj=AutoArimaForecast,test=testing)
```

```{r}
accuracy(forecast(fit1, h=24),test_ts)
```

```{r}
accuracy(forecast(fit2, h=24),test_ts)
```

```{r}
accuracy(forecast(model_auto, h=24),test_ts)
```




```{r}
#Fit the whole data
newfitwhole <- Arima(data.ts, model=model_auto)
autoplot(ts.union(data.ts,fitted(newfitwhole)),facets = FALSE)+
 scale_color_manual(labels = c("Actual", "Forecasted"),
                    values=c("black", "red")) +
 aes(linetype = series,
     size = series) +
 scale_linetype_manual(labels = c("Actual", "Forecasted"),
                       values = c(1, 4)) +
 scale_size_manual(labels = c("Actual", "Forecasted"),
                   values = c(2, 2))
```


```{r}
#Fit the whole data
newfitwhole <- Arima(data.ts, model=fit2)
autoplot(ts.union(data.ts,fitted(newfitwhole)),facets = FALSE)+
 scale_color_manual(labels = c("Actual", "Forecasted"),
                    values=c("black", "red")) +
 aes(linetype = series,
     size = series) +
 scale_linetype_manual(labels = c("Actual", "Forecasted"),
                       values = c(1, 4)) +
 scale_size_manual(labels = c("Actual", "Forecasted"),
                   values = c(2, 2))
```


